---
title: "High-Dimensional Feature Selection via Permutation Test and Bootstrap Stability Analysis"
author: "Renfei Wu, Shiqi Tian, Chengxi Li, Liwen Yin"
date: "`r Sys.Date()`"
output: pdf_document
bibliography: references.bib
nocite: "@*"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("R/data_simulation.R")
source("R/permutation_test.R")
source("R/bootstrap_stability.R")
source("R/evaluation.R")
```

## 1. Data Simulation

### (1). `simulate_mixture_data()`

This function generates a synthetic dataset for testing feature selection
methods. It creates an $n \times p$ feature matrix X** and binary labels
**Y**, where only `n_relevant` features are truly associated with Y —
the rest are pure noise. The indices of the true relevant features are
stored in **S**, which serves as ground truth for evaluation later [@R].
```{r}
sim <- simulate_mixture_data(
  n          = 200,
  p          = 50,
  n_relevant = 10,
  alpha      = 0.7,
  dist_G     = "normal",
  dist_H     = "normal",
  seed       = 123
)

dim(sim$X)   # 200 x 50
sort(sim$S)  # true relevant feature indices
table(sim$Y) # class balance
names(sim) 

```

The dataset has 200 observations and 50 features, of which 10 are truly
relevant. These indices in `sim$S` will be used as ground truth throughout
the rest of this vignette.





## 2. Permutation Testing

### (1).`stat_mean_diff()`

This function computes the absolute difference between the group means
of a single feature. A larger value indicates a stronger difference
between the two classes. which is,

$$\text{stat} = |\bar{x}_1 - \bar{x}_0|$$
```{r}
# get the most relevant feature
stats        <- sapply(sim$S, function(j) stat_mean_diff(sim$X[, j], sim$Y))
best_j       <- sim$S[which.max(stats)]

x_relevant   <- sim$X[, best_j]  # most strongly relevant feature
x_irrelevant <- sim$X[, setdiff(1:50, sim$S)[1]]   # returns all feature indices that are not contained in sim$S. Taking [1] selects the first of these indices.

stat_mean_diff(x_relevant, sim$Y)
stat_mean_diff(x_irrelevant, sim$Y)


# display plot

boxplot(x_relevant ~ sim$Y,
        main = "Relevant Feature by Class",
        xlab = "Class",
        ylab = "Value",
        col = c("lightblue", "salmon"))

boxplot(x_irrelevant ~ sim$Y,
        main = "Irrelevant Feature by Class",
        xlab = "Class",
        ylab = "Value",
        col = c("lightblue", "salmon"))
```

The relevant feature produces a larger statistic than the irrelevant
one, confirming that the function captures the signal correctly.


The boxplots show that the relevant feature exhibits a clear separation between the two classes, whereas the irrelevant feature shows substantial overlap. This visual difference aligns with the larger test statistic observed earlier.



### (2). `stat_ks()`

This function computes the Kolmogorov-Smirnov statistic, which measures
the maximum difference between the empirical distributions of a feature
in class 0 and class 1. Unlike `stat_mean_diff()`, it captures
differences in the overall shape of the distribution, not just the mean.
```{r}
stat_ks(x_relevant, sim$Y)
stat_ks(x_irrelevant, sim$Y)
```

Again, the relevant feature produces a larger statistic, indicating a
greater distributional difference between the two classes.



### (3). `stat_cvm()`

This function computes the Cramér-von Mises statistic, which measures
the overall difference between two empirical distributions by summing
the squared differences at all observed points. It is more sensitive
to differences across the whole distribution compared to `stat_ks()` [@good2005permutation].

```{r}
stat_cvm(x_relevant, sim$Y)
stat_cvm(x_irrelevant, sim$Y)
```

As expected, the relevant feature yields a larger statistic than the
irrelevant feature.


### (4). `permutation_pvalue_one()`

This function computes a permutation p-value for a single feature.
It repeatedly shuffles the class labels **Y** and recomputes the test
statistic each time, building a null distribution. The p-value is the
proportion of shuffles that produced a statistic as large as the
observed one. A small p-value suggests the feature is truly associated
with **Y**.
```{r}
set.seed(380)
permutation_pvalue_one(x_relevant, sim$Y, 
                       stat_fun   = stat_mean_diff, 
                       n_shuffles = 199)

permutation_pvalue_one(x_irrelevant, sim$Y, 
                       stat_fun   = stat_mean_diff, 
                       n_shuffles = 199)
```

The relevant feature gives a much smaller p-value than the irrelevant
feature, indicating strong evidence of association with **Y**.


### (5). `permutation_pvalues()`

This function applies `permutation_pvalue_one()` to every column of
**X**, returning a vector of p-values of length $p$. This allows us
to assess all features at once.
```{r}
set.seed(380)
pvals <- permutation_pvalues(sim$X, sim$Y,
                             stat_fun = stat_mean_diff,
                             n_shuffles = 199)
length(pvals)
summary(pvals)
```

The function returns one p-value per feature. Features with small
p-values are candidates for selection.


### (6). `select_features_permutation()`

This function selects features whose permutation p-value is at or below
a chosen significance level `alpha_level`. It is the final step of the
permutation testing pipeline.
```{r}
set.seed(380)
selected_perm <- select_features_permutation(
  sim$X, sim$Y,
  stat_fun    = stat_mean_diff,
  n_shuffles  = 199,
  alpha_level = 0.05
)

selected_perm  # selected feature indices
sim$S          # true relevant features for comparison
```

display the p-values for all 50 features:
```{r}
plot(pvals,
     xlab = "Feature index",
     ylab = "Permutation p-value",
     main = "Permutation p-values for all features",
     pch  = 16,
     col  = ifelse(1:50 %in% sim$S, "salmon", "lightblue"))
abline(h = 0.05, lty = 2)
legend("topright",
       legend = c("True signal", "Null", "alpha = 0.05"),
       col    = c("salmon", "lightblue", "black"),
       pch    = c(16, 16, NA),
       lty    = c(NA, NA, 2))
```

Features below the dashed line are selected. True signal features are
shown in red.



## 3. Bootstrap Stability

### (1). `bootstrap_selection_frequency()`

A single permutation test can vary depending on the sample. This
function runs the feature selection procedure across $B$ bootstrap
resamples and records how often each feature is selected. A higher
frequency indicates a more stable and reliable feature [@efron1994introduction].

```{r}
selector <- function(X, Y) {
  select_features_permutation(X, Y,
                              stat_fun    = stat_mean_diff,
                              n_shuffles  = 99,
                              alpha_level = 0.05)
}

set.seed(380)
freq <- bootstrap_selection_frequency(
  sim$X, sim$Y,
  select_fun = selector,
  B          = 50,
  seed       = 380
)

length(freq)
round(freq, 2)
```
Features in the true signal set tend to have higher selection
frequencies than noise features.


### (2). `select_features_bootstrap()`

This function selects features whose bootstrap selection frequency
is at or above a `threshold`. Only features that are consistently
selected across bootstrap resamples are retained, making the
selection more stable. It also returns the full frequency vector
as an attribute.
```{r}
set.seed(380)
selected_boot <- select_features_bootstrap(
  sim$X, sim$Y,
  select_fun = selector,
  B          = 50,
  threshold  = 0.5,
  seed       = 380
)

selected_boot
attr(selected_boot, "frequency")[selected_boot]

barplot(attr(selected_boot, "frequency"),
        names.arg = 1:50,
        col       = ifelse(1:50 %in% sim$S, "salmon", "lightblue"),
        xlab      = "Feature index",
        ylab      = "Selection frequency",
        main      = "Bootstrap selection frequency",
        las       = 2,
        cex.names = 0.6)
abline(h = 0.5, lty = 2)
legend("topright",
       legend = c("True signal", "Null"),
       fill   = c("salmon", "lightblue"))
```

`r length(selected_boot)` features above the dashed line are selected. True signal features
are shown in red.

## 4. Evaluation

### (1). `evaluate_precision_recall()`

This function compares the selected features against the true
relevant set **S**. Precision measures how many selected features
are truly relevant. Recall measures how many true relevant features
were found.
```{r}
pr <- evaluate_precision_recall(
  S_truth = sim$S,
  S_hat   = selected_boot
)

pr$precision
pr$recall
```

A higher precision means fewer false positives were selected. 
A higher recall means more true relevant features were found.

### (2). `f1_score()`

The F1 score combines precision and recall into a single number.
Both must be high for a good F1 score [@voss2013introduction].

$$F1 = \frac{2 \times \text{precision} \times \text{recall}}{\text{precision} + \text{recall}}$$
```{r}
f1_score(pr$precision, pr$recall)
```

An F1 score closer to 1 indicates better overall feature selection
performance.

## Generative AI Statement

Generative AI (Cursor) was used in this project to assist with two aspects: 
(1) generating the plotting code for the visualizations in the vignette, and 
(2) refining the written descriptions of each function to improve clarity 
and structure. All code and descriptions were reviewed and verified by the author.

## Bibliography


